Tail Indices and Scale Parameters in Financial Time Series

### 2017-03-05 papers/FX/draft.tex
All that can be said in the univariate case has probably been said.

Routledge and Zin \cite{routledge2010generalized} assumed
a power-law utility function of $C$:
\begin{equation}
  \label{eq:power_utility}
  u(C)=\frac{1}{1-\gamma}C^{1-\gamma}%  
\end{equation}
where $\gamma > 0$. For our discussion it suffices that $u(\cdot)$ is
an increasing function. For simplicity we assume the investor
initially has one unit of wealth. Let $r > 0$ be the return on the
risk free state bond and $X$ be the return on equity. Portfolio weights
are $\left(1-\phi\right)$ and $\phi$ respectively. Then
\begin{equation}
  \label{eq:xxie1}
  %% u(C)=\frac{1}{1-\gamma}V^{1-\gamma} \max \left\{
  %% 0, 1 + (1-\phi) r  +\phi X
  %% \right\}^{1-\gamma} \label{3}%
  C = (1 - \phi) e^r + \phi e^X
\end{equation}
For the distribution of $X$, call it $F(\cdot)$, we assume
%% \subsubsection[in case gamma > 1]{in case $\gamma > 1$}
% \subsection{power-law utility function}
% \subsubsection[in case gamma > 1]{in case $\gamma > 1$}
% In this section we consider power utility functions of
% \eqref{eq:power_utility} with $\gamma > 1$. For convenience,
% let $\xi = \gamma - 1$. In this case, we assume the following
% distribution function $F(\cdot)$ of the equity return:
\begin{equation}
  \label{eq:X_distr}
  F(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0
  \end{array}
  \right.
\end{equation}
Assume $\alpha, \beta > 1$, $0 < p < 1$ and $K > 0$.
Equation \eqref{eq:xxie0} can be re-arranged as
\begin{eqnarray}
u(v) &=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) \P[C < \delta v] \nonumber \\
&=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q) \label{eq:functional}
\end{eqnarray}
where
\begin{eqnarray*}
  q(r, \phi) &:=& \ln \left( {
    \delta v - (1 - \phi) e^r
    \over
    \phi
} \right) = \ln\left(
e^r + {\delta v - e^r \over \phi}
\right)
\end{eqnarray*}
Note $C < \delta v$ implies
\[
X < q(r, \phi)
\]
Naturally, if an agent invests in a risky asset instead of a riskless
bond, he expects to obtain a higher return from the risky asset than
he is guarranteed from the riskless bond. In our notation,
this means $\delta v > e^r$, which implies $q(r, \phi) > 0$.
The right-hand-side of \eqref{eq:functional} is a functional of the
distribution function of $X$ via \eqref{eq:xxie1}. For convenience
we denote it $\mathcal U(\cdot)$, i.e.
\[
\mathcal U F (\alpha, \beta, K, K', \phi)
= 
\E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q(r, \phi))
\]
Hereafter we abbreviate $q(r, \phi)$ to just $q$. We have
\begin{eqnarray}
  && \mathcal U F (\alpha, \beta, K, K', \phi) \nonumber \\
  &=&
  \alpha K^\alpha  p
  \int_{-\infty}^0
  u\left[ (1 - \phi) e^r + \phi e^x \right]
  {1 + b \over (K - x)^{\alpha + 1}} dx
  \nonumber \\
  &&
  \beta K'^\beta (1 - p)
  \int_{0}^\infty
  u\left[ (1 - \phi) e^r + \phi e^x \right]
  {1 + b \1{x < q} \over (K' + x)^{\beta + 1}} dx \nonumber \\
  &&
  - b u(\delta v) \left[
    1 - (1 - p) {
      K'^\beta
      \over
      (K' + q)^\beta
    }
  \right]
  \label{eq:xxie1.0}
\end{eqnarray}
Let
\[
G(\alpha, \beta, K, K') =
\max_{0 < \phi \leq 1} \mathcal U F(\alpha, \beta, K, K', \phi)
\]

\begin{theorem} \label{thrm:I}
  Let the disappointment threshold $C < \delta v$ be equivalent to $X
  < q$. Assume $q > 0$. Also assume $u(\cdot)$ is an increasing function.
  Then
  \[
  \td{G}{K} < 0 < \td{G}{\alpha}
  \]
\end{theorem}

@@

  Let
    \label{eq:xxie2}
    \hat \phi := \argmax_{0 < \phi \leq 1}
    \mathcal U F(\alpha, \beta, K, K', \phi)
  We have

@@

  G(\alpha, \beta, K, K') = \mathcal U F(\alpha, \beta, K, K', \hat \phi)

@@

  It follows
    \td{G}{\alpha}
    \pd{\mathcal U F}{\alpha}(\alpha, \beta, K, K', \hat \phi)
    + \pd{\mathcal U F}{\phi}(\alpha, \beta, K, K', \hat \phi)
    \pd{\hat \phi}{\alpha}(\alpha, \beta, K, K')
    \label{eq:xxie3}
  The definition \eqref{eq:xxie2} implies $\forall \alpha > 1$
    \label{eq:xxie4}
    \pd{\mathcal U F}{\phi}(\alpha, \beta, K, K', \hat \phi) = 0
  So the second term of \eqref{eq:xxie3} vanishes. It remains to show

@@

    \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi)
    &>& 0

@@

  We have

@@

      && \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) \\
      - \frac{\partial}{\partial \alpha} \left\{
        {\alpha p K^\alpha \over \xi} (1 + b)
        \int_{-\infty}^{0}
          \left[
            (1 - \hat\phi) e^r + \hat\phi e^x
          \right]^{-\xi}
          (K - x)^{-\alpha - 1}
        dx \right\} \\
      {p \over \xi}
      (1 + b)
      \int_{0}^{\infty}
      \left[
        (1 - \hat\phi) e^r + \hat\phi e^{-x}
      \right]^{-\xi}
      \left[
        \alpha \ln \left(
          1 + {x \over K}
        \right) - 1
      {K^{\alpha} dx
        (K + x)^{\alpha + 1}
      } \\
      - {p \over \xi}
      (1 + b)
      \int_{0}^{h}
      \left[
        (1 - \hat\phi) e^r + \hat\phi e^{-x}
      \right]^{-\xi}
      \left[
        1 - \alpha \ln \left(
          1 + {x \over K}
      {K^{\alpha} dx
        (K + x)^{\alpha + 1}
      } \\
      + {p \over \xi}
      (1 + b)
      \int_{h}^{\infty}
      \left[
        (1 - \hat\phi) e^r + \hat\phi e^{-x}
      \right]^{-\xi}
      \left[
        \alpha \ln \left(
          1 + {x \over K}
        \right) - 1
      {K^{\alpha} dx
        (K + x)^{\alpha + 1}
      } \\

@@

      - {p \over \xi}
      (1 + b)
      \left[
        (1 - \hat\phi) e^r + \hat\phi e^{-h}
      \right]^{-\xi}
      \int_{0}^{h}
      \left[
        1 - \alpha \ln \left(
          1 + {x \over K}
      {K^{\alpha} dx
        (K + x)^{\alpha + 1}
      } \\
      + {p \over \xi}
      (1 + b)
      \left[
        (1 - \hat\phi) e^r + \hat\phi e^{-h}
      \right]^{-\xi}
      \int_{h}^{\infty}
      \left[
        \alpha \ln \left(
          1 + {x \over K}
        \right) - 1
      {K^{\alpha} dx
        (K + x)^{\alpha + 1}
      } \\
      {p \over \xi}
      (1 + b)
      \left[
        (1 - \hat\phi) e^r + \hat\phi e^{-h}
      \right]^{-\xi}
      \int_{0}^{\infty}
      \left[
        \alpha \ln \left(
          1 + {x \over K}
        \right) - 1
      {K^{\alpha} dx
        (K + x)^{\alpha + 1}
      } \\
      &=& 0

@@

@@

    h = (e^{1/\alpha} - 1)K

@@

    As for $\td{G(\alpha, \beta, K, K')}{K}$, by the same argument as for
    $\td{G(\alpha, \beta, K, K')}{\alpha}$, it suffices to show

@@

    \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat\phi) < 0

@@

    We have
      && \pd{\mathcal UF}{K}(\alpha, \beta, K, K', \hat\phi) \\
      -{\alpha p (1+ b) \over \xi}
      \int_0^\infty
      \left[
        (1 - \hat\phi) e^r + \hat\phi e^{-x}
      \right]^{-\xi}
      {
        K^{\alpha - 1} (\alpha x - K)
        (K + x)^{\alpha + 2}
      dx \\
      -{\alpha p (1+ b) \over \xi}
      \left\{
        - \int_{0}^{K/\alpha}
        \left[
          (1 - \hat\phi) e^r + \hat\phi e^{-x}
        \right]^{-\xi}
        {
          K^{\alpha - 1} (K - \alpha x)
          (K + x)^{\alpha + 2}
        dx \right. \\
      \left.
        +
        \int_{K/\alpha}^{\infty}
        \left[
          (1 - \hat\phi) e^r + \hat\phi e^{-x}
        \right]^{-\xi}
        {
          K^{\alpha - 1} (\alpha x - K)
          (K + x)^{\alpha + 2}
        dx
      \right\} \\
      &<&
      -{\alpha p (1+ b) \over \xi}      
      \left[
        (1 - \hat\phi) e^r + \hat\phi e^{-M/\alpha}
      \right]^{-\xi} \times \\
      \left\{
        \int_{K/\alpha}^{\infty}
        {
          K^{\alpha - 1} (\alpha x - K)
          (K + x)^{\alpha + 2}
        dx
        - \int_{0}^{K/\alpha}
        {
          K^{\alpha - 1} (K - \alpha x)
          (K + x)^{\alpha + 2}
        dx
      \right\} \\
      &=& 0
  \end{proof}
Since $u(v)$ given in \eqref{eq:xxie1.0} as a
functional of the distribution function of $X$ is increasing with the
lower tail index $\alpha$ and decreasing with $K$, there is a curve of
equal preference on the plane of $(\alpha, K)$. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e. the estimated values of $\alpha$ and $K$
should appear positively correlated. This is indeed the case in some
real stock return data, e.g. the ``Energy'' and ``Information Technology''
sectors of S\&P 500 companies. This is illustrated in figures
\ref{fig:Energy_alpha_K} and \ref{fig:Information_Technology_alpha_K}.

% all agents guided by this class of utility
% functions will seek the largest $\alpha$ and the smallest $K$ found in
% a stock group e.g. a sector.

@@

\section{Hill Estimator}
There have been several estimators for the tail index proposed in the
literature, e.g. Hill (1975) \cite{hill1975simple}, Pikands (1975)
\cite{pickands1975statistical}, Dekkers and Einmahl and De Haan (1989)
\cite{dekkers1989moment}. See also Embrechts and Kl\"uppelberg and
Mikosch (1997)\cite{Embrechts1997}. The most popular of these is
arguably Hill (1975) \cite{hill1975simple}:
\[
\hat \alpha = \left[
  {1 \over k} \sum_{i=1}^k \ln {x_{(i)} \over x_{(k+1)}}
  \right]^{-1}
\]
where $x_{(i)}$ is the $i$-th upper order statistic of a sample
$x_1, x_2, ..., x_n$. It has been shown that the Hill estimator is
applicable to 
\begin{itemize}
\item iid random variables (Mason (1982) \cite{mason1982laws})
\item a weakly dependent sequence (Rootz{\'e}n, Leadbetter,
  and de Haan \cite{rootzen1992tail})
\item a linear process (Resnick and St{\u{a}}ric{\u{a}} (1995)
  \cite{resnick1995consistency}).
\end{itemize}
In these cases it is shown that, if $k \to \infty$ and $k/n \to 0$
as $n \to \infty$, the Hill estimator is weakly consistent, i.e.
\[
\sqrt k (\hat \alpha - \alpha) \overset{d}{\to} N(0, \alpha^2)
\]

>>>>

Routledge and Zin \cite{routledge2010generalized} assumed
a power-law utility function of $C$:
  \label{eq:power_utility}
  u(C)=\frac{1}{1-\gamma}C^{1-\gamma}%  
where $\gamma > 0$. For our discussion it suffices that $u(\cdot)$ is
an increasing function. For simplicity we assume the investor
initially has one unit of wealth. Let $r > 0$ be the return on the
risk free state bond and $X$ be the return on equity. Portfolio weights
are $\left(1-\phi\right)$ and $\phi$ respectively. Then
  \label{eq:xxie1}
  %% u(C)=\frac{1}{1-\gamma}V^{1-\gamma} \max \left\{
  %% 0, 1 + (1-\phi) r  +\phi X
  %% \right\}^{1-\gamma} \label{3}%
  C = (1 - \phi) e^r + \phi e^X
For the distribution of $X$, call it $F(\cdot)$, we assume
%% \subsubsection[in case gamma > 1]{in case $\gamma > 1$}
% \subsection{power-law utility function}
% \subsubsection[in case gamma > 1]{in case $\gamma > 1$}
% In this section we consider power utility functions of
% \eqref{eq:power_utility} with $\gamma > 1$. For convenience,
% let $\xi = \gamma - 1$. In this case, we assume the following
% distribution function $F(\cdot)$ of the equity return:
  \label{eq:X_distr}
  F(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0
  \end{array}
  \right.
Assume $\alpha, \beta > 1$, $0 < p < 1$ and $K > 0$.
Equation \eqref{eq:xxie0} can be re-arranged as
u(v) &=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) \P[C < \delta v] \nonumber \\
&=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q) \label{eq:functional}
  q(r, \phi) &:=& \ln \left( {
    \delta v - (1 - \phi) e^r
    \phi
} \right) = \ln\left(
e^r + {\delta v - e^r \over \phi}
Note $C < \delta v$ implies
X < q(r, \phi)
Naturally, if an agent invests in a risky asset instead of a riskless
bond, he expects to obtain a higher return from the risky asset than
he is guarranteed from the riskless bond. In our notation,
this means $\delta v > e^r$, which implies $q(r, \phi) > 0$.
The right-hand-side of \eqref{eq:functional} is a functional of the
distribution function of $X$ via \eqref{eq:xxie1}. For convenience
we denote it $\mathcal U(\cdot)$, i.e.
\mathcal U F (\alpha, \beta, K, K', \phi)
= 
\E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q(r, \phi))
Hereafter we abbreviate $q(r, \phi)$ to just $q$. We have
  && \mathcal U F (\alpha, \beta, K, K', \phi) \nonumber \\
  \alpha K^\alpha  p
  \int_{-\infty}^0
  u\left[ (1 - \phi) e^r + \phi e^x \right]
  {1 + b \over (K - x)^{\alpha + 1}} dx
  \nonumber \\
  \beta K'^\beta (1 - p)
  \int_{0}^\infty
  u\left[ (1 - \phi) e^r + \phi e^x \right]
  {1 + b \1{x < q} \over (K' + x)^{\beta + 1}} dx \nonumber \\
  - b u(\delta v) \left[
    1 - (1 - p) {
      K'^\beta
      (K' + q)^\beta
  \label{eq:xxie1.0}
In theorem \ref{thrm:I} we show
\max_{0 < \phi \leq 1} \mathcal U F(\alpha, \beta, K, K', \phi)
is monotonically increasing with $\alpha$ and monotonically decreasing
with $K$. But before that it is worthwhile to state the following
simple observation:
\begin{lemma} \label{lemma:I}
  Assume distribution function $F(x; \theta)$ parameterized by
  $\theta$ is defined on $x \in (a, b)$, $a, b \in \mathbb R \cup
  \{-\infty, \infty\}$, and in addition $F(x; \theta)$ has density
  function $f(x; \theta)$. Assume function $h(\cdot)$ is defined on
  $(a, b)$ and $X \sim F$.
  \begin{enumerate}
  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that
    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(x_0, b)$, 
    \pd{\E h(X)}{\theta} > 0
  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$,
    \pd{\E h(X)}{\theta} > 0
  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that
    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(x_0, b)$
    \pd{\E h(X)}{\theta} < 0
  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that 
    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$
    \pd{\E h(X)}{\theta} < 0
  \end{enumerate}
\end{lemma}

@@

  Clearly
    && \pd{\E h(X)}{\theta} \\
    &=& \int_a^b h(x) \pd{f}{\theta}(x, \theta) dx \\
    &=& \underbrace{\int_a^{x_0} h(x) \pd{f}{\theta}(x, \theta) dx}_{I_1}
    + \underbrace{\int_{x_0}^b h(x) \pd{f}{\theta}(x, \theta) dx}_{I_2}
  \begin{enumerate}
  \item When $h(x)$ is decreasing on $(a, b)$ and $\pd{f}{\theta}(x,
    \theta) > 0$ on $(a, x_0)$

@@

    I_1 > h(x_0) \int_a^{x_0} \pd{f}{\theta}(x, \theta) dx

@@

    Similarly, because $\pd{f}{\theta}(x, \theta) < 0$ for $x \in (x_0, b)$ and
    $h(x)$ is decreasing, we have

@@

      I_2 &=& \int_{x_0}^b -h(x_0)
      \left|\pd{f}{\theta}(x, \theta) \right| dx \\
      &>& -h(x_0)
      \int_{x_0}^b \left| 
        \pd{f}{\theta}(x, \theta)
      \right| dx

@@

    Finally we have

@@

      && \pd{\E h(X)}{\theta} \\
      &=& I_1 + I_2 \\
      &>& h(x_0) \int_a^b \pd{f}{\theta}(x, \theta) dx \\
      &=& h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx \\
      &=& h(x_0) \opd{\theta} 1 = 0
  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$
      I_1 &=&
      \int_a^{x_0} -h(x)
      \left| \pd{f}{\theta}(x, \theta) \right| dx \\

@@

      -h(x_0) \int_a^{x_0}
      \left| \pd{f}{\theta}(x, \theta) \right| dx

@@

    and obviously

@@

    I_2 > h(x_0) \int_{x_0}^b
    \pd{f}{\theta}(x, \theta) dx

@@

    So

@@

    \pd{\E h(X)}{\theta}
    = I_1 + I_2
    > h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx = 0

@@

  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that
    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(x_0, b)$

    Obviously
    \[
    I_1 = \int_a^{x_0} h(x) \pd{f}{\theta}(x, \theta) dx
    < h(x_0) \int_a^{x_0} \pd{f}{\theta}(x, \theta) dx
    \]
    and
    \begin{eqnarray*}
      I_2 &=& \int_{x_0}^b h(x) \pd{f}{\theta}(x, \theta) dx \\
      &<& -h(x_0) \int_{x_0}^b \left| \pd{f}{\theta}(x, \theta) \right| dx \\
    \end{eqnarray*}
    Thus
    \[
    \pd{\E h(X)}{\theta}
    = I_1 + I_2
    < h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx = 0
    \]
  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that 
    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$

    In this case
    \begin{eqnarray*}
      I_1 &=& \int_a^{x_0} h(x) \pd{f}{\theta}(x, \theta) dx \\
      &<& -h(x_0) \int_a^{x_0} \left| \pd{f}{\theta}(x, \theta) \right| dx
    \end{eqnarray*}
    And
    \begin{eqnarray*}
      I_2 &=& \int_{x_0}^b h(x) \pd{f}{\theta}(x, \theta) dx \\
      &<& h(x_0) \int_{x_0}^b \pd{f}{\theta}(x, \theta) dx
    \end{eqnarray*}
    So
    \[
    \pd{\E h(X)}{\theta}
    = I_1 + I_2
    < h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx = 0
    \]
  \end{enumerate}
\end{proof}
With lemma \ref{lemma:I} it is straightforward to show the
monotonicity claimed earlier.
Let
\[
G(\alpha, \beta, K, K') =
\max_{0 < \phi \leq 1} \mathcal U F(\alpha, \beta, K, K', \phi)
\]
\begin{theorem}
  \label{thrm:I}
  \[
  \pd{G}{\alpha} (\alpha, \beta, K, K') > 0,
  \quad
  \pd{G}{K} (\alpha, \beta, K, K') < 0
  \]
\end{theorem}

\begin{proof}
  Let
  \begin{equation}
    \label{eq:xxie2}
    \hat \phi := \argmax_{0 < \phi \leq 1}
    \mathcal U F(\alpha, \beta, K, K', \phi)
  \end{equation}
  We have
  \[
  G(\alpha, \beta, K, K') = \mathcal U F(\alpha, \beta, K, K', \hat \phi)
  \]
  It follows
  \begin{eqnarray}
    \pd{G}{\alpha}
    &=&
    \pd{\mathcal U F}{\alpha}(\alpha, \beta, K, K', \hat \phi)
    + \pd{\mathcal U F}{\phi}(\alpha, \beta, K, K', \hat \phi)
    \pd{\hat \phi}{\alpha}(\alpha, \beta, K, K')
    \label{eq:xxie3}
  \end{eqnarray}
  The definition \eqref{eq:xxie2} implies $\forall \alpha > 1$
  \begin{equation}
    \label{eq:xxie4}
    \pd{\mathcal U F}{\phi}(\alpha, \beta, K, K', \hat \phi) = 0
  \end{equation}
  So the second term of \eqref{eq:xxie3} vanishes. It remains to show
  \begin{eqnarray*}
    \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi)
    &>& 0
  \end{eqnarray*}
  From \eqref{eq:xxie1.0}, it follows
  \begin{eqnarray*}
    && \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat\phi)
    \\
    &=& p \opd{\alpha}\E[u((1 - \phi) e^r + \phi e^X) | X < 0]
  \end{eqnarray*}
  The function $u((1 - \phi) e^r + \phi e^X)$ is obviously increasing
  with $X$. Furthermore
  \begin{eqnarray*}
    && \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
    &=&
    - {K^\alpha \over (K - x)^{\alpha + 1}}
    \left[
      \alpha
      \ln\left(
        1 - {x \over K}
      \right) - 1
    \right]
  \end{eqnarray*}
  It is easily checked that when $x < K(1 - e^{1/\alpha}) < 0$,
  \[
  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
  \]
  and when $K(1 - e^{1/\alpha}) < x < 0$,
  \[
  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
  \]
  This is the second case of lemma \ref{lemma:I}. So we have
  \begin{eqnarray*}
    \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) &>& 0
  \end{eqnarray*}
  As for $\pd{G(\alpha, \beta, K, K')}{K}$, by the same argument as for
  $\pd{G(\alpha, \beta, K, K')}{\alpha}$, it suffices to show
  \[
  \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat\phi) < 0
  \]
  We have
  \[
  \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat\phi)
  = p \opd{K} \E \left[
    u((1 - \phi) e^r + \phi e^X)
    | X < 0
  \right] 
  \]
  and
  \begin{eqnarray*}
    && \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
    &=&
    -\alpha K^{\alpha - 1} {
      \alpha x + K
      \over
      (K - x)^{\alpha + 2}
    }
  \end{eqnarray*}
  Clearly, when $x < -\alpha/K$
  \[
  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
  \]
  and when $-\alpha/K < x < 0$
  \[
  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
  \]
  So by the third case of lemma \ref{lemma:I} we conclude
  \[
  \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat\phi) < 0  
  \]
\end{proof}
Since $u(v)$ given in \eqref{eq:xxie1.0} as a
functional of the distribution function of $X$ is increasing with the
lower tail index $\alpha$ and decreasing with $K$, there is a curve of
equal preference on the plane of $(\alpha, K)$. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e. the estimated values of $\alpha$ and $K$
should appear positively correlated. This is indeed the case in some
real stock return data, e.g. the ``Energy'' and ``Information Technology''
sectors of S\&P 500 companies. This is illustrated in figures
\ref{fig:Energy_alpha_K} and \ref{fig:Information_Technology_alpha_K}.
The confidence bands of the Hill estimates of $\alpha$ of these
sectors are shown in figures \ref{fig:Energy_lower},
\ref{fig:Consumer_Staples_lower} and \ref{fig:Information_Technology_lower}.
% all agents guided by this class of utility
% functions will seek the largest $\alpha$ and the smallest $K$ found in
% a stock group e.g. a sector.

@@

\section{Hill Estimator}
\label{sec:Hill}
There have been several estimators for the tail index proposed in the
literature, e.g. Hill (1975) \cite{hill1975simple}, Pikands (1975)
\cite{pickands1975statistical}, Dekkers and Einmahl and De Haan (1989)
\cite{dekkers1989moment}. See also Embrechts and Kl\"uppelberg and
Mikosch (1997)\cite{Embrechts1997}. The most popular of these is
arguably Hill (1975) \cite{hill1975simple}:
\[
\hat \alpha = \left[
  {1 \over k} \sum_{i=1}^k \ln {x_{(i)} \over x_{(k+1)}}
  \right]^{-1}
\]
where $x_{(i)}$ is the $i$-th upper order statistic of a sample
$x_1, x_2, ..., x_n$. It has been shown that the Hill estimator is
applicable to 
\begin{itemize}
\item iid random variables (Mason (1982) \cite{mason1982laws})
\item a weakly dependent sequence (Rootz{\'e}n, Leadbetter,
  and de Haan \cite{rootzen1992tail})
\item a linear process (Resnick and St{\u{a}}ric{\u{a}} (1995)
  \cite{resnick1995consistency}).
\end{itemize}
In these cases it is shown that, if $k \to \infty$ and $k/n \to 0$
as $n \to \infty$, the Hill estimator is weakly consistent, i.e.
\[
\sqrt k (\hat \alpha - \alpha) \overset{d}{\to} N(0, \alpha^2)
\]
